Probabilistic Forecast Reconciliation

### 2018-11-06 Presentation/ProbForeRec.tex
background to point forecast reconciliati rework

\documentclass{beamer}
\usetheme{Warsaw}
\usepackage{amsmath,amsthm, bm, tikz,qtree}
\usetikzlibrary{arrows,positioning,shapes,fit,calc}
\title{Probabilistic Forecast Reconciliation}
\date{August 15, 2018}
\author[GPAH]{Puwasala Gamakumara, Anastasios Panagiotelis, George Athanasopoulos and Rob Hyndman}
\begin{document}
  \begin{frame}
    \maketitle
  \end{frame}
  \section{Concepts}
  \begin{frame}{Motivating Examples}
  \begin{itemize}
  	\item Gross Domestic Product
  	\begin{itemize}
  		\item An aggregate of consumption, investment, government spending and trade balance
  		\item Further breakdown, e.g. consumption of food, rent, etc.
  	\end{itemize}
    \item Retail sales
    \begin{itemize}
    	\item Aggregate of different product category sales
    	\item Further breakdown into individual stock keeping units.  
    \end{itemize}
    \item Potentially need forecasts of all time series.
  \end{itemize}
  \end{frame}
  \begin{frame}{Incoherent Forecasts}
  \begin{itemize}
    \item Potential approaches
  	\begin{itemize}
  		\item Univariate models
  		\item Multivariate models
  		\item Judgemental forecasts
  	\end{itemize}
  \item Forecasts may not respect aggregation structure. ({\bf Incoherent})
  \item Outcome {\em will} respect aggregation  structure. ({\bf Coherent})
  \item Motivation is aggregation but can be generalised to any linear constraints.
  \end{itemize}
  \end{frame}
  \begin{frame}{Reconciliation}
  \begin{itemize}
  	\item Begin with a vector of forecasts that are incoherent.
  	\item Adjust these {\bf ex post} to make them coherent.
  	\item There are good solutions for point forecasting that:
  	\begin{itemize}
  		\item Guarantee coherent forecasts.
  		\item Improve forecast accuracy overall.
  	\end{itemize}
  	\item Generalisation to probabilistic forecasts is our contribution.
  	\item Getting there necessitates a rethink of the existing point forecasting literature. 
  \end{itemize}
  \end{frame}

  \begin{frame}{Hierarchical and Grouped Time Series}

@@

  	  \item Collections of time series are often characterised by aggregation constraints

@@

  	    \item Cross-Sectionally
  		\item Temporally

@@

      \item {\bf Coherent} forecasts respect such constraints. 
      \item Independently produced forecasts are generally incoherent.

@@

@@

		\leaf{AA} \leaf{AB} 

@@

		\leaf{BA} \leaf{BB}

@@

  \begin{frame}{Forecast Reconciliation}

@@

  	  \item Forecast reconciliation involves
  	  	\item Producing incoherent {\bf base} forecasts for all series in an $n\times 1$ vector $\hat{\bm y}$
        \item Adjusting base forecasts to obtain coherent {\bf reconciled} forecasts in an $n\times 1$ vector $\tilde{\bm y}$
      \item Why do we care?
        	\item Aligned decision making.
        	\item Improved forecast accuracy

@@

  \begin{frame}{Reconciliation in two steps}
  		\item Many reconciliation methods involve two steps
  			\item Pre-multiply $\hat{\bm y}$ by a $m\times n$ matrix $\bm G$ to obtain {\bf bottom} level series ${\bm b}={\bm G}{\hat{\bm y}}$ 
  			\item Pre-multiply ${\bm b}$ by a $n\times m$ matrix $\bm S$ to obtain ${\tilde{\bm y}}$, i.e. $\tilde{\bm y}={\bm S}{{\bm b}}$
  	    \item The matrix ${\bm S}$ defines the aggregation constraints, e.g.

@@

  	    \item Choice of ${\bm G}$ defines reconciliation method, e.g. OLS: ${\bm G}=\left(\bm{S}'\bm{S}\right)^{-1}{\bm S'}$ and Bottom Up: ${\bm G}=\left(\bm{0}_{m\times n-m}~\bm{I}_{m\times m}\right)$ 

@@

@@

    Instead of using bottom-level series a different combination of $m$ {\bf basis series} could be used (e.g. top and $m-1$ bottom).  Although ${\bm S}$ would be different $\mathfrak{s}$ would be the same.

@@

  \begin{frame}{Special Case: Projection}
   \item An important special case is where ${\bm S}{\bm G}$ is a projection.
  	\item ${\bm S}{\bm G}$ is symmetric
  	\item ${\bm S}{\bm G}$ is idempotent
  \item Let ${\bm v}\in\mathfrak{s}$ 
    	\item ${\bm S}{\bm G}{\bm v}$ will also lie in $\mathfrak{s}$.
    	\item ${\bm S}{\bm G}{\bm v}={\bm v}$ only when ${\bm S}{\bm G}$ is a projection.

@@

@@

  \begin{frame}{Projections preserve unbiasedness}
	Let $\hat{\bm y}_{t+h|t}$ be an unbiased forecast that is $E_{1:t}(\hat{\bm y}_{t+h|t})={\bm \mu_{t+h|t}}$ where ${\bm \mu_{t+h|t}}=E(\bm{y}_{t+h}\mid\bm{y}_{1},\dots,\bm{y}_{t})$
	  The reconciled forecast $\tilde{\bm y}_{t+h|t}={\bm S}{\bm G}\hat{\bm y}_{t+h|t}$ will also be unbiased iff ${\bm S}{\bm G}$ is a projection.
    Previously, this was often stated as an assumption that ${\bm S}{\bm G}{\bm S}={\bm S}$.    
  \begin{frame}{Proof}
  Very easy proof
  E_{1:t}(\tilde{\bm{y}}_{t+h|t})
  &= E_{1:t}(\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t})\\
  &= \bm{S}\bm{G}E_{1:t}(\hat{\bm{y}}_{t+h|t})\\
  &= \bm{S}\bm{G}\bm{\mu}_{t+h|t}\\
  &= \bm{\mu}_{t+h|t}
  The last equality does not hold for ${\bm G}$ in general but does hold when ${\bm S}{\bm G}$ is a projection.

@@

  	\item Oblique projections have the same property for non-Euclidean distance. 

>>>>

\documentclass{beamer}
\usetheme{Warsaw}
\usepackage{amsmath,amsthm, bm, tikz,qtree}
\usetikzlibrary{arrows,positioning,shapes,fit,calc}
\usepackage[greek,english]{babel}

\title{Probabilistic Forecast Reconciliation}
\date{August 15, 2018}
\author[GPAH]{Puwasala Gamakumara, Anastasios Panagiotelis, George Athanasopoulos and Rob Hyndman}
\begin{document}
  \begin{frame}
    \maketitle
  \section{Motivation}
  \begin{frame}{Motivating Examples}

@@

  	\item Gross Domestic Product

@@

  		\item An aggregate of consumption, investment, government spending and trade balance
  		\item Further breakdown, e.g. consumption of food, rent, etc.

@@

    \item Retail sales
    	\item Aggregate of different product category sales
    	\item Further breakdown into individual stock keeping units.  

@@

    \item Potentially need forecasts of all time series.
  \begin{frame}{Incoherent Forecasts}
    \item Potential approaches
  		\item Univariate models
  		\item Multivariate models
  		\item Judgemental forecasts
  \item Forecasts do not respect aggregation structure ({\bf Incoherent})
  \item Outcome {\em does} respect aggregation  structure ({\bf Coherent})
  \item Motivation is aggregation but can be generalised to any linear constraints.
  \begin{frame}{Reconciliation}
  	\item Begin with a vector of forecasts that are incoherent.
  	\item Adjust these {\bf ex post} to make them coherent.
  	\item There are good solutions for point forecasting that:
  		\item Guarantee coherent forecasts.
  		\item Improve forecast accuracy overall.
  	\item Generalisation to probabilistic forecasts is our contribution.
  	\item Getting there necessitates a rethink of the existing point forecasting literature. 
  \begin{frame}{What we do NOT do}
  	\item All information is contained in the most disaggregate series.
  	\item In principle using the correct multivariate model for the most disaggregate series and aggregating them should work.
  	\item Disaggregate series are: 
  		\item Very noisy
  		\item High-dimensional
  		\item Prone to model misspecification
  \section{Point Forecast Reconciliation}
 \begin{frame}{A simple hierarchy}
  	Consider a hierarchy given by

@@

		\leaf{\alt<2>{\color{blue}AA}{AA}} \leaf{\alt<2>{\color{blue}AB}{AB}} 

@@

		\leaf{\alt<2>{\color{blue}BA}{BA}} \leaf{\alt<2>{\color{blue}BB}{BB}}

@@

@@

    	\item Let $n$ be the number of series, $\bm{y}$ be an n-vector of all series.
    	\only<2>{\item Let {\color{blue}$m$} be the bottom level series and {\color{blue}$\bm{b}$} be an {\color{blue}$m$}-vector of the bottom level series.}

@@

\begin{frame}{The ${\bm S}$ matrix}
    The matrix ${\bm S}$ defines the aggregation constraints, e.g.

@@

    Coherence holds when
    {\bm y}={\bm S}{\bm b}
\begin{frame}{As a regression model}
  	\item Cast the problem as a regression model with base forecasts $\hat{\bm{y}}$ as the ``dependent variable'' and ${\bm S}$ as the ``design matrix''. 
  	\hat{\bm{y}}={\bm{S}}{\bm\beta}+{\bm e}
  	\item Initial approach (Athanasopoulos et al, 2009; Hyndman et al, 2011) was to fit by OLS  yielding reconciled forecasts:
  	\tilde{\bm{y}}={\bm S}({\bm S}'{\bm S})^{-1}{\bm S}'{\hat{\bm{y}}}

@@

\begin{frame}{Generalisation}
  \item Wherever we can use OLS we can use GLS
  \tilde{\bm{y}}={\bm S}({\bm S}'{\bm W}^{-1}{\bm S})^{-1}{\bm S}'{\bm W}^{-1}{\hat{\bm{y}}}
  \item Diagonal ${\bm W}$ considered by Athanasopoulos et al (2017)
  \item MinT approach (Wickremasuriya et al, 2018) use a ${\bm W}$ that is an estimate of the {\em in-sample} forecast error covariance matrix.
%  \begin{frame}{Reconciliation in two steps}
%  	\begin{itemize}
%  		\item Many reconciliation methods involve two steps
%  		\begin{enumerate}
%  			\item Pre-multiply $\hat{\bm y}$ by a $m\times n$ matrix $\bm G$ to obtain {\bf bottom} level series ${\bm b}={\bm G}{\hat{\bm y}}$ 
%  			\item Pre-multiply ${\bm b}$ by a $n\times m$ matrix $\bm S$ to obtain ${\tilde{\bm y}}$, i.e. $\tilde{\bm y}={\bm S}{{\bm b}}$
%  		\end{enumerate}
%
%  	    \item Choice of ${\bm G}$ defines reconciliation method, e.g. OLS: ${\bm G}=\left(\bm{S}'\bm{S}\right)^{-1}{\bm S'}$ and Bottom Up: ${\bm G}=\left(\bm{0}_{m\times n-m}~\bm{I}_{m\times m}\right)$ 
%    \end{itemize}
%  \end{frame}
\begin{frame}
\includegraphics[height=5cm]{Figs/Plato.jpeg}\\
\textgreek{AGEWMTETRHTOS MHDEIS EISITW}\\
Those without knowledge of geometry may not enter.

@@

    Instead of using bottom-level series a different combination of $m$ {\bf basis series} could be used (e.g. top and $m-1$ bottom).\\
    Although ${\bm S}$ would be different $\mathfrak{s}$ would be the same.

@@

@@

%  \begin{frame}{Special Case: Projection}
%  \begin{itemize}
%   \item An important special case is where ${\bm S}{\bm G}$ is a projection.
%  \begin{itemize}
%  	\item ${\bm S}{\bm G}$ is symmetric
%  	\item ${\bm S}{\bm G}$ is idempotent
%  \end{itemize}
%  \item Let ${\bm v}\in\mathfrak{s}$ 
%    \begin{itemize}
%    	\item ${\bm S}{\bm G}{\bm v}$ will also lie in $\mathfrak{s}$.
%    	\item ${\bm S}{\bm G}{\bm v}={\bm v}$ only when ${\bm S}{\bm G}$ is a projection.
%    \end{itemize}
%  \end{itemize}
%  \end{frame}
%
%  \begin{frame}{Geometry: Oblique Projection}
%  	\vspace{-0.9cm}
%  	\centering
%  	\input{Figs/pointforerec_schematic}
%  \end{frame}
%  \begin{frame}{Projections preserve unbiasedness}
%	Let $\hat{\bm y}_{t+h|t}$ be an unbiased forecast that is $E_{1:t}(\hat{\bm y}_{t+h|t})={\bm \mu_{t+h|t}}$ where ${\bm \mu_{t+h|t}}=E(\bm{y}_{t+h}\mid\bm{y}_{1},\dots,\bm{y}_{t})$
%	\begin{theorem}
%	  The reconciled forecast $\tilde{\bm y}_{t+h|t}={\bm S}{\bm G}\hat{\bm y}_{t+h|t}$ will also be unbiased iff ${\bm S}{\bm G}$ is a projection.
%    \end{theorem}
%    Previously, this was often stated as an assumption that ${\bm S}{\bm G}{\bm S}={\bm S}$.    
%  \end{frame}
%  \begin{frame}{Proof}
%  Very easy proof
%  \begin{align*}
%  E_{1:t}(\tilde{\bm{y}}_{t+h|t})
%  &= E_{1:t}(\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t})\\
%  &= \bm{S}\bm{G}E_{1:t}(\hat{\bm{y}}_{t+h|t})\\
%  &= \bm{S}\bm{G}\bm{\mu}_{t+h|t}\\
%  &= \bm{\mu}_{t+h|t}
%  \end{align*}
%  The last equality does not hold for ${\bm G}$ in general but does hold when ${\bm S}{\bm G}$ is a projection.
%  \end{frame}
  	\item What about the MinT approach? 
  \begin{frame}{Finding a direction}
  \item Consider the covariance matrix of ${\bm y}-\hat{\bm{y}}$.
   \item This can be estimated using in-sample forecast errors.
  \item This provides information about the likely direction of an error.
  \item Projecting along this direction is more likely to result in a reconciled forecast closer to the target.
  \begin{frame}{Animation}
  content...

@@

  \begin{frame}{Is this overkill?}

@@

  	\item Is this geometric interpretation really necessary?
  	\item When we generalise to probabilistic forecasts the regression interpretation does not really fit.
  	\item Geometric ideas can easily be generalised. 
